Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-19 paper/paper.tex
Revised section 5

\begin{figure}[tb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
\caption{Pareto-like distribution of workflow sizes in ensemble. }
\label{fig:ensemble-distribution}
\end{figure}

@@

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:number-complete-pareto}
\end{figure*}

\begin{figure}[b!] 
\centering
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of algorithms based on score, which  is computed as
$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
priority).}
\label{fig:score}
\end{figure}

In order to evaluate the algorithms on a standard ser of workflows, we used
ensembles constructed from the DAGs available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi08}. The gallery contains
workflows of five applications from Pegasus and from Indiana
University~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
workflows of different sizes based on a given distribution. Two distributions
are considered:
\begin{itemize}
  \item ensemble of workflows of equal size (constant)
  \item Pareto-like distribution.
\end{itemize}

@@

\begin{figure*}[t]  
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{ Total work completed expressed as the sum of runtimes (in hours) of
tasks of completed workflows.Data computed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:total-time}
\end{figure*}

@@

\begin{equation}
\label{eq:score}
s = \sum_{DAGs\ completed}(1/(p+1))
\end{equation}

where priorities of subsequent workflows in the ensemble are $p=0,1,2,\ldots$ (0
means highest priority). This scoring function gives more weight to the top high
priority workflows, whereas the low priority ones are trated almost equally.

@@

\begin{figure*}[htb] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Effective computing cost in \$ per hour calclated by dividing sum of
task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows.}
\label{fig:cost}
\end{figure*}

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values per each application ensemble. The budget values
selected correspond to two opposite cases: {\em small budget} (top) and {\em
large budget} (bottom). The first one allows completing only a small number of
workflows from the ensemble, while the latter one suffices to finish almost all
the workflows. It can be seen that the static planning algorithm (SPSS) performs
better for tighter constraints: when the budget is low to admit only a small
number of workflows, the static procedure is able to more efficiently pack these
workflow tasks onto the resources. This can be clearly seen for the LIGO,
Epigenomics and SIPHT ensembles for small budgets and also for the same
ensembles with large budgets but short deadlines. The exception are Montage and
CyberShake ensembles, which consist of many fine-grained tasks that can be
executed in parallel. For these cases, as well for the other ensembles with
large budgets, the simple dynamic scheduling techniqies perform equally well or
better. The reason for that behavior is that the large number of tasks
guarantees that resources are almost never idle, while the provisioner algorithm
makes sure that the VM instances are most efficiently utilized before shutdown.
We can also observe that the workflow-aware dynamic algorithm (WA-DPDS) performs
better that the workflow-unaware version. This means that for the tested
workflows the simple admission procedure based on estimation of workflow cost
does not degrade the solution, but rather in many cases it allows rejecting
larger workflows that would lead to budget overrun. Thus it can save the space
for smaller workflows that can complete.

Fig.~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
computed for two sample applications. Comparing
Fig.~\ref{fig:number-complete-pareto} to Fig.~\ref{fig:score} we can see that
the latter one gives more flattened shape: this means that e.g. the improvement
of the solution by one workflow may mean adding a one low-priority workflow.
Nevertheless we can see that the curves on both figures have a very similar
shape, which mens that our algorithms obey the priority rules and do not try to
admit lower-priority workflows at the cost of the more important ones.

@@

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric $T_C$. These plots confirm our previous observations about the
algorithms performance, but they also allow to observe a ore general property.
For a given budget, when the deadline is tightened the number of work done
steeply decreases. Assiging a shorter deadline means that more VMs need to be
provisioned to complete the work, but this leads to higher parallelism and thus
to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
costs are important, parallelism should be avoided when possible.

These observations are further confirmed when examining the efffective resource
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increase of actual costs of computing. This increase reaches ca.
20\%, which means that if we need to run e.g. a SIPHT workflow ensemble with a
short deadline, ca. \$200 out of \$1000 is spent on idle resources.

Another general concluson can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We can see that the selection of deadline and cost can be formulated as
multi-criteria optimization problem and the curves presented in
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the number of work done, whcich direcly results in minimization of effective cost parameter for a given
deadline. Therefore, the obtained results can be used to assist the decison in
resource allocation planning, when both cost and time criteria are given not as
constraints but as objectives.

% \begin{figure*}[htb] 
% \centering
% \includegraphics[width=0.19\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
% \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
% Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
% priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
% priority).}
% \label{fig:}
% \end{figure*}

>>>>

\begin{figure}[tb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
\caption{Pareto-like distribution of workflow sizes in ensemble. Size is
measured in number of tasks.}
\label{fig:ensemble-distribution}
\end{figure}

@@

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows, {\em small budget}
(top) and {\em large budget} (bottom).}
\label{fig:number-complete-pareto}
\end{figure*}

%\begin{figure}[b!] 
%\centering
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-GENOME-n-1000-8-dagh100-1500m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
%\caption{Comparison of algorithms based on score, which  is computed as
%$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$
%(0 means highest priority).}
%\label{fig:score}
%\end{figure}

In order to evaluate the algorithms on a standard ser of workflows, we used
ensembles constructed from the DAGs available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi08}. The gallery contains
workflows of five applications from Pegasus and from Indiana
University~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
workflows of different sizes based on a given distribution. By the size of the
workflow means the number of jobs. Two distributions are considered:
\begin{itemize}
  \item ensemble of workflows of equal size (constant)
  \item Pareto-like distribution.
\end{itemize}

@@

\begin{figure*}[t]  
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{ Total work completed expressed as the sum of runtimes (in hours) of
tasks of completed workflows. Data computed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:total-time}
\end{figure*}

@@

%\begin{equation}
%\label{eq:score}
%s = \sum_{DAGs\ completed}(1/(p+1))
%\end{equation}

\begin{equation}
\label{eq:score}
s = \sum_{DAGs\ completed}2^{-p}
\end{equation}

where priorities of subsequent workflows in the ensemble are $p=0,1,2,\ldots$ (0
means highest priority). This exponential scoring function gives a strong weight
to the top priority workflows, since it has a property that a score for
completing a single workflow with priority $p$ is higher that the score for
completing all lower priority ones: 
\begin{equation}
\label{eq:score-property}
2^{-p} > \sum_{i=p+1,\ldots}2^{-i}
\end{equation}

@@

\begin{figure*}[htb] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Effective computing cost in \$ per hour calculated by dividing sum of
task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows.}
\label{fig:cost}
\end{figure*}

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values per each application ensemble. The budget values
selected correspond to two opposite cases: {\em small budget} (top) and {\em
large budget} (bottom). The first one allows completing only a small number of
workflows from the ensemble, while the latter one suffices to finish almost all
the workflows. It can be seen that the static planning algorithm (SPSS) performs
better for tighter constraints: when the budget is low to admit only a small
number of workflows, the static procedure is able to more efficiently pack these
workflow tasks onto the resources. This can be clearly seen for the LIGO,
Epigenomics and SIPHT ensembles for small budgets and also for the same
ensembles with large budgets but short deadlines. The exception are Montage and
CyberShake ensembles, which consist of many fine-grained tasks that can be
executed in parallel. For these cases, as well for the other ensembles with
large budgets, the simple dynamic scheduling techniques perform equally well or
better. The reason for that behavior is that the large number of tasks
guarantees that resources are almost never idle, while the provisioner algorithm
makes sure that the VM instances are most efficiently utilized before shutdown.
We can also observe that the workflow-aware dynamic algorithm (WA-DPDS) performs
better than the workflow-unaware version. This means that for the tested
workflows the simple admission procedure based on estimation of workflow cost
does not degrade the solution, but rather in many cases it allows rejecting
larger workflows that would lead to budget overrun. Thus it can save the space
for smaller workflows that can complete.

%Fig.~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
%computed for two sample applications. Comparing
%Fig.~\ref{fig:number-complete-pareto} to Fig.~\ref{fig:score} we can see that
%the latter one gives more flattened shape: this means that e.g. the improvement
%of the solution by one workflow may mean adding a one low-priority workflow.
%Nevertheless we can see that the curves on both figures have a very similar
%shape, which mens that our algorithms obey the priority rules and do not try to
%admit lower-priority workflows at the cost of the more important ones.

@@

\begin{table}[tb]
\centering

(a) Number of workflows completed $N_c$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & TOTAL\tabularnewline
\hline
DPDS      &   57  & 45 &  18 &  14  &  0 & 134\tabularnewline
WA-DPDS   &    96 &  87  & 48  & 34  &  0 & 265\tabularnewline
SPSS     &    29  & 42  & 192  & 74  & 50 & 387\tabularnewline
\end{tabular}
\medskip

(b) Total amount of computing time completed $T_C$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & TOTAL\tabularnewline
\hline
DPDS      &   58  & 45 &  19 &  14  &  0 & 136\tabularnewline
WA-DPDS   &    90  & 96  & 57  & 27 &   8 & 278\tabularnewline
SPSS     &    31 &  18 &  170  & 75 &  42 & 336\tabularnewline
\end{tabular}
\medskip

(c) Exponential score $s$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & TOTAL\tabularnewline
\hline
DPDS      &   58 &  45 &  46 &  15  &  9 & 173\tabularnewline
WA-DPDS   &    90 &  96  & 84  & 29 &  17 & 316\tabularnewline
SPSS     &    48 &  27 & 173 &  75 &  42 & 365\tabularnewline
\end{tabular}
\medskip

M -- Montage; C -- CyberShake; L -- LIGO; E -- Epigenomics; S --SIPHT

\label{tab:num-dags-pareto}
\caption{Number of times each algorithm achieved the highest score for each
application (over all budgets and deadlines). Data based on total 525 runs on Pareto-distributed ensembles.}
\end{table}

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric $T_C$. These plots confirm our previous observations about the
algorithms performance, but they also allow to observe a more general property.
For a given budget, when the deadline is tightened the number of work done
steeply decreases. Assiging a shorter deadline means that more VMs need to be
provisioned to complete the work, but this leads to higher parallelism and thus
to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
costs are important, parallelism should be avoided when possible.

These observations are further confirmed when examining the efffective resource
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increase of actual costs of computing. This increase reaches
$\sim$20\%, which means that if we need to run e.g. a SIPHT workflow ensemble
with a short deadline, $\sim$\$200 out of \$1000 is spent on idle resources.

Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We can see that the selection of deadline and cost can be formulated as
multi-criteria optimization problem and the curves presented in
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the number of work done, which direcly results in minimization of effective cost parameter for a given
deadline. Therefore, the obtained results can be used to assist the decison in
resource allocation planning, when both cost and time criteria are given not as
constraints but as objectives.

% \begin{figure*}[htb] \centering
% \includegraphics[width=0.19\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
% \includegraphics[width=0.19\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
% \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
% Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
% priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
% priority).} \label{fig:} \end{figure*}

In order to compare efficiency of algorithms, we counted the number of times
each algorithm achieved the best results for all tested applications, deadlines
and budgets. If two or more algorithms achieve the same highest results, all are
counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values.
These results depend on the metric: number of workflows completed does not count
the priorities or sizes of workflows, which means that it happens more often
that all the algorithms achieve the same (best) result. However, it does not
mean that all algorithms complete the same workflows: that is why values
in the tables (a), (b) and (c) are different. it can be observed that based on
the sampled problem space the workflow-aware algorithms achieve highest score
in $\sim$2--3 times more cases than the simple DPDS which does not take into
account workflow structure or size. In average, SPSS performs better in
$\sim$20--50\% more times than WA-DPDS. We have to note, that the sampled
paramter space included the Montage and CyberShake workflows consisting of many
small tasks, for which cases SPSS does not perform equally well.
